### 2017-07-07 paper/app.tex
added example for runtime bound

>>>>

\section{Computations for the proof of Theorem~\ref{THM:RUNTIME-BOUND}}

We 
prove, by induction on $k$, that for each $k$ it holds 
$\E[\nodecrank{k}{j}]\leq 
\vec{c}[j]\cdot\left(\sum_{j'<j}\E[\nodecrank{k}{j'}] \right)+ 
\E[\vecseq{X}{0}{j}] - \E[\vecseq{X}{k}{j}]$.

The base case $k=0$ is simple as both sides of the inequality are zero. Assume 
that the inequality holds for some $k\geq 0$. We have 
$\E[\nodecrank{k+1}{j}]=\E[\nodecrank{k}{j}]+\probm{(L_j^{k})}$, so from 
induction hypothesis we get 
\begin{equation}
\label{eq:time1}
\E[\nodecrank{k+1}{j}]\leq 
\vec{c}[j]\cdot\left(\sum_{j'<j}\E[\nodecrank{k}{j'}] \right)+ 
\E[\vecseq{X}{0}{j}] - \E[\vecseq{X}{k}{j}] + \probm{(L_j^{k})}.\end{equation} 
Now denote 
$L^k_{<j} = L^k_1 \cup \dots\cup L^k_{j-1}$ and $L^k_{>j}= 
L^k_{j+1}\cup\dots\cup L^k_{d}$. We have  
$\E[\vecseq{X}{k}{j}] = \E[\vecseq{X}{k}{j}\cdot 
\indicator{ L_{<j}^k}] + \E[\vecseq{X}{k}{j}\cdot 
\indicator{ L_j^k}] + \E[\vecseq{X}{k}{j}\cdot 
	\indicator{L_{>j}^k}] \geq 
\E[\vecseq{X}{k+1}{j}\cdot 
	\indicator{ L_{<j}^k}] -\vec{c}[j]\cdot \probm(L_{<j}^k) + 
	\E[\vecseq{X}{k+1}{j}\cdot 
	\indicator{ L_{j}^k}] + \probm(L_{j}^k) + \E[\vecseq{X}{k+1}{j}\cdot 
	\indicator{ L_{>j}^k}]= \E[\vecseq{X}{k+1}{j}] -\vec{c}[j]\cdot 
	\probm(L_{<j}^k)+ \probm(L_{j}^k)$. Plugging this 
	into~\ref{eq:time1} yields
\begin{align*}
\E[\nodecrank{k+1}{j}]&\leq 
\vec{c}[j]\cdot\left(\sum_{j'<j}\E[\nodecrank{k}{j'}] \right)+ 
\E[\vecseq{X}{0}{j}] - \E[\vecseq{X}{k+1}{j}] + \vec{c}[j]\cdot 
\probm(L_{<j}^k)\\
&=\vec{c}[j]\cdot\left(\sum_{j'<j}\E[\nodecrank{k+1}{j'}] \right)+ 
\E[\vecseq{X}{0}{j}] - \E[\vecseq{X}{k+1}{j}].
\end{align*}

### 2017-07-07 paper/compositional.tex
added example for runtime bound

\lstset{language=affprob}
\lstset{tabsize=2}
\newsavebox{\unintthree}
\begin{lrbox}{\unintthree}
	\begin{lstlisting}[mathescape]
while $x\geq 0$ do
	while $y\geq 0$ do
		$z:=x$;
		while $z\geq 0$ do
			$z:=z-1$;
			$x:=x-1$
		od;
		$y:=y-1$
	od;
	$x:=x+$sample$(\mathit{Uniform[-3,1]})$
od
	\end{lstlisting}
\end{lrbox}
\newsavebox{\unintfour}
\begin{lrbox}{\unintfour}
	\begin{lstlisting}[mathescape]
while $x\geq 0 $ do
	$y:=x$;
	while $y\geq 1$ do
		$y:=y+$sample$(\mathit{Uniform[-3,1]})$
	od;
	$x:=x-1$
od
	\end{lstlisting}
\end{lrbox}
\begin{figure}[t]
	%%\centering
	\begin{subfigure}{0.45\textwidth}
		\usebox{\unintfour}
		\vspace{0.2cm}
		\begin{tikzpicture}[x=1.8cm,y=1.4cm]
		\node[det] (init) at (0,0) {$\loc_0$};
		\node[det] (fin) at (0,-1) {$\loc^{\lout}$};
		\node[det] (as1) at (1,0) {$\loc_1$};
		\node[det] (inner) at (2,0) {$\loc_2$};
		\node[det] (as2) at (2,-1) {$\loc_3$};
		\node[det] (as3) at (3,0) {$\loc_4$};
%		
		\draw[tran] (init) to node[font=\scriptsize,draw, fill=white, 
		rectangle,pos=0.5] {$x<0$} (fin);
		\draw[tran] (init) to node[font=\scriptsize,draw, fill=white, 
		rectangle,pos=0.5] {$x\geq 0$} (as1);
		\draw[tran] (as1) -- node[auto,font=\scriptsize] {$y:=x$} (inner);
		\draw[tran] (inner) to node[font=\scriptsize,draw, fill=white, 
		rectangle,pos=0.5] {$y<1$} (as3);
		\draw[tran] (inner) to node[font=\scriptsize,draw, fill=white, 
		rectangle,pos=0.5] {$y\geq 1$} (as2);
		\draw[tran,rounded corners] (as2) -- (1.25,-1) -- 
		node[font=\scriptsize, 
		label={[xshift=-0.4cm,yshift=-0.2cm,font=\scriptsize] 
		{$y:=\dots$}}] 
		{} 
		(inner);
		\draw[tran, rounded corners] (as3) -- (3,0.5) -- node[font=\scriptsize, 
		label={[font=\scriptsize, yshift=-0.1cm] {$x:=x-1$}}] {} (0,0.5) -- 
		(init);
		\end{tikzpicture}
		%	\hspace{1cm}
		%	\usebox{\uniintt}
		\caption{Example for slicing illustration: program and its pCFG.}
		\label{fig:uniint4}
	\end{subfigure}
	\hspace{0.5cm}
	\begin{subfigure}{0.45\textwidth}
	\usebox{\unintthree}
	%	\hspace{1cm}
	%	\usebox{\uniintt}
	\caption{Program where the outer loop does not have a PVSM that 
		typechecks.}
	\label{fig:uniint2}
	\end{subfigure}
\caption{Program illustrations.}
\label{fig:comp-ex}
\end{figure}

@@

The soundness of the algorithm follows from Theorems~\ref{thm:nonneg-comp} and 
\ref{thm:holger-comp}. Note that we use the PVSM-based algorithm 
of~\cite{HolgerPOPL} as a back-up sub-procedure for the case when LinNCSM-based 
proof fails. Hence, Algorithm~\ref{algo:compositional} can compositionally 
prove a.s. termination of strictly larger class of programs than the PVSM-based 
algorithm alone.

>>>>

\lstset{language=affprob}
\lstset{tabsize=2}
\newsavebox{\unintthree}
\begin{lrbox}{\unintthree}
	\begin{lstlisting}[mathescape]
while $x\geq 0$ do
	while $y\geq 0$ do
		$z:=x$;
		while $z\geq 0$ do
			$z:=z-1$;
			$x:=x-1$
		od;
		$y:=y-1$
	od;
	$x:=x+$sample$(\mathit{Uniform[-3,1]})$
od
	\end{lstlisting}
\end{lrbox}
\newsavebox{\unintfour}
\begin{lrbox}{\unintfour}
	\begin{lstlisting}[mathescape]
while $x\geq 0$ and $z \geq 0 $ do
	$y:=z$;
	while $y\geq 3$ do
		$y:=y+$sample$(\mathit{Uniform[-3,1]})$
	od;
	$x:=x-1$
od
	\end{lstlisting}
\end{lrbox}
\begin{figure}[t]
	%%\centering
	\begin{subfigure}{0.45\textwidth}
		\usebox{\unintfour}
		\vspace{0.2cm}
		\begin{tikzpicture}[x=1.8cm,y=1.4cm]
		\node[det] (init) at (0,0) {$\loc_0$};
		\node[det] (fin) at (0,-1) {$\loc^{\lout}$};
		\node[det] (as1) at (1.2,0) {$\loc_1$};
		\node[det] (inner) at (2,0) {$\loc_2$};
		\node[det] (as2) at (2,-1) {$\loc_3$};
		\node[det] (as3) at (3,0) {$\loc_4$};
%		
		\draw[tran] (init) to node[font=\scriptsize,draw, fill=white, 
		rectangle,pos=0.6] {$x<0 \vee z<0$} (fin);
		\draw[tran] (init) to node[font=\scriptsize,draw, fill=white, 
		rectangle,pos=0.5,text width=0.8cm] {$x\geq 0$\\$\wedge$\\$z\geq 0$} 
		(as1);
		\draw[tran] (as1) -- node[auto,font=\scriptsize] {$y:=z$} (inner);
		\draw[tran] (inner) to node[font=\scriptsize,draw, fill=white, 
		rectangle,pos=0.5] {$y<3$} (as3);
		\draw[tran] (inner) to node[font=\scriptsize,draw, fill=white, 
		rectangle,pos=0.5] {$y\geq 3$} (as2);
		\draw[tran,rounded corners] (as2) -- (1.25,-1) -- 
		node[font=\scriptsize, 
		label={[xshift=-0.4cm,yshift=-0.2cm,font=\scriptsize] 
		{$y:=\dots$}}] 
		{} 
		(inner);
		\draw[tran, rounded corners] (as3) -- (3,0.5) -- node[font=\scriptsize, 
		label={[font=\scriptsize, yshift=-0.1cm] {$x:=x-1$}}] {} (0,0.5) -- 
		(init);
		\end{tikzpicture}
		%	\hspace{1cm}
		%	\usebox{\uniintt}
		\caption{Example for slicing illustration: program and its pCFG.}
		\label{fig:uniint4}
	\end{subfigure}
	\hspace{0.5cm}
	\begin{subfigure}{0.45\textwidth}
	\usebox{\unintthree}
	%	\hspace{1cm}
	%	\usebox{\uniintt}
	\caption{Program where the outer loop does not have a PVSM that 
		typechecks.}
	\label{fig:uniint2}
	\end{subfigure}
\caption{Program illustrations.}
\label{fig:comp-ex}
\end{figure}

@@

The soundness of the algorithm follows from Theorems~\ref{thm:nonneg-comp} and 
\ref{thm:holger-comp}. Note that we can use the PVSM-based algorithm 
of~\cite{HolgerPOPL} as a back-up sub-procedure for the case when LinNCSM-based 
proof fails. Hence, Algorithm~\ref{algo:compositional} can compositionally 
prove a.s. termination of strictly larger class of programs than the PVSM-based 
algorithm alone.

### 2017-07-07 paper/extensions.tex
added example for runtime bound

\begin{theorem}
\label{thm:runtime-bound}
Let $\{\vec{X}_{i}\}_{i=0}^{\infty}$ be an 
$n$-dimensional strict LexRSM with $c$-bounded ECI for some stopping time $\stime$. 
Then $\E[\stime]\leq  
\sum_{j=1}^{n}\E[\vecseq{X}{0}{j}]\cdot(\vec{c}[j]+1)^{n-j}$.
\end{theorem}
\begin{proof}
Fix an instance $(\{\vec{X}_{i=0}^{\infty},\{L_1^i,\dots,L_{n+1}^i\}_{i=0}^{\infty})$ satisfying Definition~\ref{def:lexrsm-eci}. Denote $\noofdecrank_j(\omega)$ the number of steps $i$ in which $\omega\in 
L_j^i$. Since by Theorem~\ref{thm:lexrsm-main} the existence of strict LexRSM entails 
$\probm(\stime<\infty)=1$, the value $\noofdecrank_j(\omega)$ is a.s. finite for all $1\leq j \leq n$. 
We prove that for each $1\leq j \leq n$ it holds $\E[\noofdecrank_j]\leq 
\vec{c}[j]\cdot\left(\sum_{j'<j}\E[\noofdecrank_{j'}]\right) + 
\E[\vecseq{X}{0}{j}].$ Since $\stime(\omega)=\sum_{1\leq j \leq n} 
\noofdecrank_j(\omega)$, for each $\omega\in \Omega$ (and hence, due to 
linearity of expectation $\E[\stime]=\sum_{1\leq j \leq n} 
E[\noofdecrank_j]$), the statement of the 
Theorem follows by an easy induction.

To prove the required inequality, let $\nodecrank{k}{j}(\omega)$ be the number 
of steps $i$ \emph{within the first $k$} steps such that $\omega\in L_j^i$. We 
prove, by induction on $k$, that for each $k$ it holds 
$\E[\nodecrank{k}{j}]\leq 
\vec{c}[j]\cdot\left(\sum_{j'<j}\E[\nodecrank{k}{j'}] \right)+ 
\E[\vecseq{X}{0}{j}] - \E[\vecseq{X}{k}{j}]$. Once this is proved, the 
desired inequality follows by taking $k$ to $\infty$, since 
$\lim_{k\rightarrow \infty}\E[\nodecrank{k}{j}] = \E[\noofdecrank_j]$ and 
$\lim_{k\rightarrow \infty}\E[\vecseq{X}{k}{j}] \geq 0$.

The base case $k=0$ is simple as both sides of the inequality are zero. Assume 
that the inequality holds for some $k\geq 0$. We have 
$\E[\nodecrank{k+1}{j}]=\E[\nodecrank{k}{j}]+\probm{(L_j^{k})}$, so from 
induction hypothesis we get 
\begin{equation}
\label{eq:time1}
\E[\nodecrank{k+1}{j}]\leq 
\vec{c}[j]\cdot\left(\sum_{j'<j}\E[\nodecrank{k}{j'}] \right)+ 
\E[\vecseq{X}{0}{j}] - \E[\vecseq{X}{k}{j}] + \probm{(L_j^{k})}.\end{equation} 
Now denote 
$L^k_{<j} = L^k_1 \cup \dots\cup L^k_{j-1}$ and $L^k_{>j}= 
L^k_{j+1}\cup\dots\cup L^k_{d}$. We have  
$\E[\vecseq{X}{k}{j}] = \E[\vecseq{X}{k}{j}\cdot 
\indicator{ L_{<j}^k}] + \E[\vecseq{X}{k}{j}\cdot 
\indicator{ L_j^k}] + \E[\vecseq{X}{k}{j}\cdot 
	\indicator{L_{>j}^k}] \geq 
\E[\vecseq{X}{k+1}{j}\cdot 
	\indicator{ L_{<j}^k}] -\vec{c}[j]\cdot \probm(L_{<j}^k) + 
	\E[\vecseq{X}{k+1}{j}\cdot 
	\indicator{ L_{j}^k}] + \probm(L_{j}^k) + \E[\vecseq{X}{k+1}{j}\cdot 
	\indicator{ L_{>j}^k}]= \E[\vecseq{X}{k+1}{j}] -\vec{c}[j]\cdot 
	\probm(L_{<j}^k)+ \probm(L_{j}^k)$. Plugging this 
	into~\ref{eq:time1} yields
\begin{align*}
\E[\nodecrank{k+1}{j}]&\leq 
\vec{c}[j]\cdot\left(\sum_{j'<j}\E[\nodecrank{k}{j'}] \right)+ 
\E[\vecseq{X}{0}{j}] - \E[\vecseq{X}{k+1}{j}] + \vec{c}[j]\cdot 
\probm(L_{<j}^k)\\
&=\vec{c}[j]\cdot\left(\sum_{j'<j}\E[\nodecrank{k+1}{j'}] \right)+ 
\E[\vecseq{X}{0}{j}] - \E[\vecseq{X}{k+1}{j}].
\end{align*}
%
%\end{align}
\end{proof}

To transfer this mathematical result to probabilistic programs, we want to 
impose a restriction on LexRSM maps that ensures that all components of a 
LexRSM map have, from each reachable configuration, an expected one-step 
increase of at most $c$. Here $c$ can be a constant, but it can also be a value 
that depends on the initial configurations: this is to handle cases where some 
variables are periodically reset to a value related to the initial variable 
values, such as variable $z$ in Figure~\ref{fig:uniint2}. To this end, let 
$\program$ be a \PP{} with a pCFG $\pCFG_\program$ and let 
$\vec{\lem}=(\lem_1,\dots,\lem_n)$ be an 
$n$-dimensional $1$-LexRSM map for $\program$. Consider an $n$-dimensional 
vector 
$\vec{\bar{c}}=(\bar{c}_1,\dots,\bar{c}_n)$ whose each component is an 
expression over variables of 
$\program$. We say 
that $\vec{\lem}$ has 
$\vec{\bar{c}}$-bounded ECI w.r.t. invariant map $\inv$ if the following holds 
for each initial configuration $(\locinit,\vecinit)$ with $\vecinit\in 
\vecinitset$: for 
each 
configuration $(\loc,\vec{x})$ with $\vec{x}\in \inv(\loc)$ and generalized 
transition $\tilde{\tau}$ 
of 
$\pCFG_\program$ outgoing from $(\loc,\vec{x})$ it holds that if $j$ is 
the smallest index such that 
$\tilde{\tau}$ is $1$-ranked by $\lem_j$ from $(\loc,\vec{x})$, then for all 
$j'>j$ the gen. transition $\tilde{\tau}$ is $f$-ranked by $\lem_{j'}$ from 
$(\loc,\vec{x})$, where $f=-c_{j'}(\vecinit)$. From 
Theorem~\ref{thm:runtime-bound} we have the following:

\begin{corollary}
\label{col:runtime-progs}
Let $\program$ be a probabilistic program. Assume that there exists an 
$n$-dimensional $\eps$-LexRSM map $\vec{\lem}=(\lem_1,\dots,\lem_d)$ for 
$\program$ supported 
by some 
invariant map $\inv$, scuh that $\vec{\lem}$ has $\vec{\bar{c}}$-bounded ECI 
(w.r.t. $\inv$) 
for some vector of expressions $\vec{\bar{c}}=(\bar{c}_1,\dots,\bar{c}_n)$. 
Then under each scheduler $\sigma$ and for each initial valuation of program 
variables $\vecinit\in\vecinitset$ it holds $\E^{\sigma}_{\vecinit}[\ttime]\leq 
\sum_{j=1}^{n}\lem_j(\locinit,\vecinit)\cdot(\bar{c}_j(\vecinit))^{n-j}$.
\end{corollary}

>>>>

\begin{theorem}
\label{thm:runtime-bound}
\label{THM:RUNTIME-BOUND}
Let $\{\vec{X}_{i}\}_{i=0}^{\infty}$ be an 
$n$-dimensional strict LexRSM with $c$-bounded ECI for some stopping time $\stime$. 
Then $\E[\stime]\leq  
\sum_{j=1}^{n}\E[\vecseq{X}{0}{j}]\cdot(1+\sum_{j'= j+1}^{n}\prod_{j'}^{n} 
\vec{c}[j'])$.
\end{theorem}
\begin{proof}
Fix an instance $(\{\vec{X}_{i=0}^{\infty},\{L_1^i,\dots,L_{n+1}^i\}_{i=0}^{\infty})$ satisfying Definition~\ref{def:lexrsm-eci}. Denote $\noofdecrank_j(\omega)$ the number of steps $i$ in which $\omega\in 
L_j^i$. Since by Theorem~\ref{thm:lexrsm-main} the existence of strict LexRSM entails 
$\probm(\stime<\infty)=1$, the value $\noofdecrank_j(\omega)$ is a.s. finite for all $1\leq j \leq n$. 
We prove that for each $1\leq j \leq n$ it holds $\E[\noofdecrank_j]\leq 
\vec{c}[j]\cdot\left(\sum_{j'<j}\E[\noofdecrank_{j'}]\right) + 
\E[\vecseq{X}{0}{j}].$ Since $\stime(\omega)=\sum_{1\leq j \leq n} 
\noofdecrank_j(\omega)$, for each $\omega\in \Omega$ (and hence, due to 
linearity of expectation $\E[\stime]=\sum_{1\leq j \leq n} 
E[\noofdecrank_j]$), the statement of the 
Theorem follows by an easy induction.

To prove the required inequality, let $\nodecrank{k}{j}(\omega)$ be the number 
of steps $i$ \emph{within the first $k$} steps such that $\omega\in L_j^i$. We 
prove, by induction on $k$, that for each $k$ it holds 
$\E[\nodecrank{k}{j}]\leq 
\vec{c}[j]\cdot\left(\sum_{j'<j}\E[\nodecrank{k}{j'}] \right)+ 
\E[\vecseq{X}{0}{j}] - \E[\vecseq{X}{k}{j}]$. Once this is proved, the 
desired inequality follows by taking $k$ to $\infty$, since 
$\lim_{k\rightarrow \infty}\E[\nodecrank{k}{j}] = \E[\noofdecrank_j]$ and 
$\lim_{k\rightarrow \infty}\E[\vecseq{X}{k}{j}] \geq 0$. The inductive proof is 
somewhat technical and deferred to \AppendixMaterial.

%The base case $k=0$ is simple as both sides of the inequality are zero. Assume 
%that the inequality holds for some $k\geq 0$. We have 
%$\E[\nodecrank{k+1}{j}]=\E[\nodecrank{k}{j}]+\probm{(L_j^{k})}$, so from 
%induction hypothesis we get 
%\begin{equation}
%\label{eq:time1}
%\E[\nodecrank{k+1}{j}]\leq 
%\vec{c}[j]\cdot\left(\sum_{j'<j}\E[\nodecrank{k}{j'}] \right)+ 
%\E[\vecseq{X}{0}{j}] - \E[\vecseq{X}{k}{j}] + \probm{(L_j^{k})}.\end{equation} 
%Now denote 
%$L^k_{<j} = L^k_1 \cup \dots\cup L^k_{j-1}$ and $L^k_{>j}= 
%L^k_{j+1}\cup\dots\cup L^k_{d}$. We have  
%$\E[\vecseq{X}{k}{j}] = \E[\vecseq{X}{k}{j}\cdot 
%\indicator{ L_{<j}^k}] + \E[\vecseq{X}{k}{j}\cdot 
%\indicator{ L_j^k}] + \E[\vecseq{X}{k}{j}\cdot 
%	\indicator{L_{>j}^k}] \geq 
%\E[\vecseq{X}{k+1}{j}\cdot 
%	\indicator{ L_{<j}^k}] -\vec{c}[j]\cdot \probm(L_{<j}^k) + 
%	\E[\vecseq{X}{k+1}{j}\cdot 
%	\indicator{ L_{j}^k}] + \probm(L_{j}^k) + \E[\vecseq{X}{k+1}{j}\cdot 
%	\indicator{ L_{>j}^k}]= \E[\vecseq{X}{k+1}{j}] -\vec{c}[j]\cdot 
%	\probm(L_{<j}^k)+ \probm(L_{j}^k)$. Plugging this 
%	into~\ref{eq:time1} yields
%\begin{align*}
%\E[\nodecrank{k+1}{j}]&\leq 
%\vec{c}[j]\cdot\left(\sum_{j'<j}\E[\nodecrank{k}{j'}] \right)+ 
%\E[\vecseq{X}{0}{j}] - \E[\vecseq{X}{k+1}{j}] + \vec{c}[j]\cdot 
%\probm(L_{<j}^k)\\
%&=\vec{c}[j]\cdot\left(\sum_{j'<j}\E[\nodecrank{k+1}{j'}] \right)+ 
%\E[\vecseq{X}{0}{j}] - \E[\vecseq{X}{k+1}{j}].
%\end{align*}
%
%\end{align}
\end{proof}

To transfer this mathematical result to probabilistic programs, we want to 
impose a restriction on LexRSM maps that ensures that all components of a 
LexRSM map have, from each reachable configuration, an expected one-step 
increase of at most $c$. Here $c$ can be a constant, but it can also be a value 
that depends on the initial configurations: this is to handle cases where some 
variables are periodically reset to a value related to the initial variable 
values, such as variable $y$ in Figure~\ref{fig:uniint4}. To this end, let 
$\program$ be a \PP{} with a pCFG $\pCFG_\program$ and let 
$\vec{\lem}=(\lem_1,\dots,\lem_n)$ be an 
$n$-dimensional $1$-LexRSM map for $\program$. Consider an $n$-dimensional 
vector 
$\vec{\bar{c}}=(\bar{c}_1,\dots,\bar{c}_n)$ whose each component is an 
expression over variables of 
$\program$. We say 
that $\vec{\lem}$ has 
$\vec{\bar{c}}$-bounded ECI w.r.t. invariant map $\inv$ if the following holds 
for each initial configuration $(\locinit,\vecinit)$ with $\vecinit\in 
\vecinitset$: for 
each 
configuration $(\loc,\vec{x})$ with $\vec{x}\in \inv(\loc)$ and generalized 
transition $\tilde{\tau}$ 
of 
$\pCFG_\program$ outgoing from $(\loc,\vec{x})$ it holds that if $j$ is 
the smallest index such that 
$\tilde{\tau}$ is $1$-ranked by $\lem_j$ from $(\loc,\vec{x})$, then for all 
$j'>j$ the gen. transition $\tilde{\tau}$ is $f$-ranked by $\lem_{j'}$ from 
$(\loc,\vec{x})$, where $f=-c_{j'}(\vecinit)$. From 
Theorem~\ref{thm:runtime-bound} we have the following:

\begin{corollary}
\label{col:runtime-progs}
Let $\program$ be a probabilistic program. Assume that there exists an 
$n$-dimensional $\eps$-LexRSM map $\vec{\lem}=(\lem_1,\dots,\lem_d)$ for 
$\program$ supported 
by some 
invariant map $\inv$, scuh that $\vec{\lem}$ has $\vec{\bar{c}}$-bounded ECI 
(w.r.t. $\inv$) 
for some vector of expressions $\vec{\bar{c}}=(\bar{c}_1,\dots,\bar{c}_n)$. 
Then under each scheduler $\sigma$ and for each initial valuation of program 
variables $\vecinit\in\vecinitset$ it holds $\E^{\sigma}_{\vecinit}[\ttime]\leq
\sum_{j=1}^{n}\lem_j(\locinit,\vecinit)\cdot(1+\sum_{j'= j+1}^{n}\prod_{j'}^{n} 
\bar{c}_{j'}(\vecinit)).$
%\sum_{j=1}^{n}\lem_j(\locinit,\vecinit)\cdot(\bar{c}_j(\vecinit))^{n-j}$.
\end{corollary}

\begin{example}
Consider the program in Figure~\ref{fig:uniint4}. For the program we have an 
invariant map $\inv$ with $\inv(\loc_0)=x\geq -1$, $\inv(\loc_1)=\inv(\loc_4) = 
x\geq 0 
\wedge z\geq 0$, $\inv(\loc_2)=\inv(\loc_3)= x\geq 0 \wedge y \geq 0
\wedge z\geq 0$, and $\inv(\loc^{\lout})=\mathit{true}$. Further, there exists 
the following 
$3$-dimensional 
$1$-LinLexRSM map $\vec{\lem}$ supported by $\inv$: 
$\vec{\lem}(\loc_0)=(2x+3,0,2)$, $\vec{\lem}(\loc_1)=(2x+3,0,1)$, 
$\vec{\lem}(\loc_2)=(2x+2,y,1)$, $\vec{\lem}(\loc_3)=(2x+2,y,0)$, 
$\vec{\lem}(\loc_4)=(2x+2,0,0)$, and $\vec{\lem}(\loc^{\lout})=(0,0,0)$. It is 
easy to check that $\vec{\lem}$ is has $(0,z,2)$-bounded ECI. Applying 
Corollary~\ref{col:runtime-progs} yields $\E[\ttime]\leq (2x_{\mathit{init}} + 
3)\cdot(1+3z_{\mathit{init}})+0+2= 6x_{\mathit{init}}z_{\mathit{init}} + 
2x_{\mathit{init}} + 9z_{\mathit{init}}+5$. Note that this quadratic bound is 
asymptotically optimal for the given program.
\end{example}
